### 2022-09-13 main/chapter_searches.tex
motivate cls and arrange some citations

Misinterpretations of \pvalues\ are widespread and well reported~\cite{
schervish1996p,
Cohen1994TheEI,
Amrhein2017TheEI,
goodman2008dirty,
greenland2016no,
Bernstein2016princess,
wagenmakers2007practical,
Trafimow2015ban
}.
For example, the idea of ``probability of compatibility'' with a model~\cite{
HIGG-2018-04
}
appears to be spreading as a good meme~\cite{dawkins1989selfish} in
High Energy Physics~\cite{
HIGG-2018-27,
HIGG-2018-51,
EXOT-2018-08,
HION-2018-19,
HIGG-2017-09,
langford2021combination,
mastrandrea2019searches,
white2019search,
IceCube2013search,
IceCube2014searches,
gerasimov2021new
},
plausibly explained by vague but accurate statements like
``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
lhcb2021test
}
mutating to precise and inaccurate statements like
``a probability of around $0.1\%$ that the data is compatible with the
Standard Model predictions''~\cite{cern2021test}.

There also exists a notion that a small \pvalue\ is significant because
it is a rare fluctuation.
All \pvalues\ are rare ---
by construction they are uniformly distributed, so the
probability of $P(h) < \varepsilon$ is equal to the probability of
$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set
(if the data are sampled under $h$).
All that matters to the interpretation is that an alternative hypothesis makes
the observed result \emph{more probable}.

% CLs
$\cls$~\cite{
zech1988cls,
bock1998lower,
read1997optimal,
Read2002cls,
Murray2010heretic
}
\begin{equation}
\cls
=
\frac{P(s + b)}{P(b)}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}
}
.
\end{equation}
These \pvalues\ are sometimes labelled $\clsb = P(s + b)$ and $\clb = P(b)$
Some literature also defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
},
where $p_b$ is not quite a \pvalue\ if it excludes equality with the test
statistic.

@@

textbooks cited by pdg~\cite{
eadie1971statistical,
frodesen1979probability
}

>>>>

Misrepresentations of \pvalues\ (and other frequentist statistics) are
widespread and widely reported~\cite{
schervish1996p,
Cohen1994TheEI,
Amrhein2017TheEI,
goodman2008dirty,
greenland2016no,
Bernstein2016princess,
wagenmakers2007practical,
Trafimow2015ban
}.
For example, the idea of ``probability of compatibility'' with a model~\cite{
HIGG-2018-04
}
appears to be spreading as a good meme~\cite{dawkins1989selfish} in
High Energy Physics~\cite{
HIGG-2018-27,
HIGG-2018-51,
EXOT-2018-08,
HION-2018-19,
HIGG-2017-09,
langford2021combination,
mastrandrea2019searches,
white2019search,
IceCube2013search,
IceCube2014searches,
gerasimov2021new
},
as plausibly explained by mutations from vaguely accurate statements like
``compatible with the SM prediction with a $P$ value of $0.10\%$''~\cite{
lhcb2021test
}
to precisely inaccurate statements like
``a probability of around $0.1\%$ that the data is compatible with the
Standard Model predictions''~\cite{cern2021test}.

There also exists a notion that a small \pvalue\ is significant because
it is a rare fluctuation~\cite{murray1997use}.
All \pvalues\ are rare ---
by construction, they are uniformly distributed, so the
probability of $P(h) < \varepsilon$ is equal to the probability of
$0.4 \leq P(h) < 0.4 + \varepsilon$ or any other $\varepsilon$-sized set.
If, that is, the data are sampled under $h$ ---
what matters to the interpretation is whether an alternative hypothesis makes
the observed result \emph{more probable}.

For example, maximum likelihood ratios (and variations on their theme) are
useful and commonly applied, but they can give \pvalues\ that appear
pathological --- consider a Poisson background ($b$) added to an
unconstrained signal yield ($s$).
A sensible interpretation would find that `excess' data well above the
background expectation favour the background-plus-signal hypothesis ($s + b$)
---
but if the signal is allowed to be negative, then a `deficit' of data below the
background can have a maximum likelihood ratio that favours $s + b$.
This scrambles excesses and deficits together in the same tail, so both get
small \pvalues, and so a small \pvalue\ implies neither a preference for the
signal hypothesis nor incompatibility with the background.

As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
on obfuscated data, so they can be interpreted in ratios.
When setting upper limits on disfavoured signal models at the LHC,
it is standard practice to use such \pvalue\ ratios name $\cls$~\cite{
read2000modified,
Read2002cls,
Murray2010heretic,
pdg2022ynf
},
which defined as
\begin{equation}
\cls
=
\frac{P(s + b)}{P(b)\hphantom{s +\;}}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
}
.
\end{equation}
These individual \pvalues\ are sometimes labelled
$\clsb = P(s + b)$ and $\clb = P(b)$.%
\footnote{%
Despite the $\mathrm{CL}$ labels, these are not confidence levels ---
a confidence level $\gamma$ is a boundary at which \pvalues\ are selected, such
as to set limits by solving for $x$ such that
$\gamma = 1 - P(x)$~\cite{pdg1996}.
In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
at least 1984~\cite{pdg1984}, although earlier textbooks
(cited by~\cite{pdg1984}) do make the distinction~\cite{
eadie1971statistical,
frodesen1979probability
}.
}
Some literature also defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
},
but $p_b$ may not quite be a \pvalue\ if it excludes equality with the observed
test statistic.

% history: bayes uniform prior in special cases
% just an approximation
% use for Higgs at LEP to avoid spurious rejection of bkg
LEP Higgs boson searches~\cite{
read1997optimal,
bock1998lower,
etde1998prospects,
junk1999confidence,
lep2000searches
}

credible intervals in simple Poisson or Gaussian cases~\cite{
Helene1983upper,
pdg1988,
pdg2022ynf
}

origins~\cite{
pdg1988,
zech1988cls,
bock1998lower,
read1997optimal,
Read2002cls,
Murray2010heretic
}

@@
